\documentclass[12pt,letterpaper]{article}
\usepackage{graphicx,textcomp}
\usepackage{natbib}
\usepackage{setspace}
\usepackage{fullpage}
\usepackage{color}
\usepackage[reqno]{amsmath}
\usepackage{amsthm}
\usepackage{fancyvrb}
\usepackage{amssymb,enumerate}
\usepackage[all]{xy}
\usepackage{endnotes}
\usepackage{lscape}
\newtheorem{com}{Comment}
\usepackage{float}
\usepackage{hyperref}
\newtheorem{lem} {Lemma}
\newtheorem{prop}{Proposition}
\newtheorem{thm}{Theorem}
\newtheorem{defn}{Definition}
\newtheorem{cor}{Corollary}
\newtheorem{obs}{Observation}
\usepackage[compact]{titlesec}
\usepackage{dcolumn}
\usepackage{tikz}
\usetikzlibrary{arrows}
\usepackage{multirow}
\usepackage{xcolor}
\newcolumntype{.}{D{.}{.}{-1}}
\newcolumntype{d}[1]{D{.}{.}{#1}}
\definecolor{light-gray}{gray}{0.65}
\usepackage{url}
\usepackage{listings}
\usepackage{color}

\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolour}{rgb}{0.95,0.95,0.92}

\lstdefinestyle{mystyle}{
	backgroundcolor=\color{backcolour},   
	commentstyle=\color{codegreen},
	keywordstyle=\color{magenta},
	numberstyle=\tiny\color{codegray},
	stringstyle=\color{codepurple},
	basicstyle=\footnotesize,
	breakatwhitespace=false,         
	breaklines=true,                 
	captionpos=b,                    
	keepspaces=true,                 
	numbers=left,                    
	numbersep=5pt,                  
	showspaces=false,                
	showstringspaces=false,
	showtabs=false,                  
	tabsize=2
}
\lstset{style=mystyle}
\newcommand{\Sref}[1]{Section~\ref{#1}}
\newtheorem{hyp}{Hypothesis}

\title{Problem Set 2}
\date{Due: February 19, 2023}
\author{Applied Stats II}


\begin{document}
	\maketitle
	\section*{Instructions}
	\begin{itemize}
		\item Please show your work! You may lose points by simply writing in the answer. If the problem requires you to execute commands in \texttt{R}, please include the code you used to get your answers. Please also include the \texttt{.R} file that contains your code. If you are not sure if work needs to be shown for a particular problem, please ask.
		\item Your homework should be submitted electronically on GitHub in \texttt{.pdf} form.
		\item This problem set is due before 23:59 on Sunday February 19, 2023. No late assignments will be accepted.
	%	\item Total available points for this homework is 80.
	\end{itemize}

	
	%	\vspace{.25cm}
	
%\noindent In this problem set, you will run several regressions and create an add variable plot (see the lecture slides) in \texttt{R} using the \texttt{incumbents\_subset.csv} dataset. Include all of your code.

	\vspace{.25cm}
%\section*{Question 1} %(20 points)}
%\vspace{.25cm}
\noindent We're interested in what types of international environmental agreements or policies people support (\href{https://www.pnas.org/content/110/34/13763}{Bechtel and Scheve 2013)}. So, we asked 8,500 individuals whether they support a given policy, and for each participant, we vary the (1) number of countries that participate in the international agreement and (2) sanctions for not following the agreement. \\

\noindent Load in the data labeled \texttt{climateSupport.csv} on GitHub, which contains an observational study of 8,500 observations.

\begin{itemize}
	\item
	Response variable: 
	\begin{itemize}
		\item \texttt{choice}: 1 if the individual agreed with the policy; 0 if the individual did not support the policy
	\end{itemize}
	\item
	Explanatory variables: 
	\begin{itemize}
		\item
		\texttt{countries}: Number of participating countries [20 of 192; 80 of 192; 160 of 192]
		\item
		\texttt{sanctions}: Sanctions for missing emission reduction targets [None, 5\%, 15\%, and 20\% of the monthly household costs given 2\% GDP growth]
		
	\end{itemize}
	
\end{itemize}

\newpage
\noindent Please answer the following questions:

\begin{enumerate}
	\item
	Remember, we are interested in predicting the likelihood of an individual supporting a policy based on the number of countries participating and the possible sanctions for non-compliance.
	\begin{enumerate}
		\item [] Fit an additive model. Provide the summary output, the global null hypothesis, and $p$-value. Please describe the results and provide a conclusion.
		%\item
		%How many iterations did it take to find the maximum likelihood estimates?
		
		\lstinputlisting[firstline=65, lastline=70]{PS2.R}
		\begin{verbatim}
			> summary(modl)
			Call:glm(formula = choice ~ countries + sanctions, 
			family = "binomial",     data = dat)
			Deviance Residuals:     
			Min       1Q   Median       3Q      Max 
			-1.4259  -1.1480  -0.9444   1.1505   1.4298  
			Coefficients:            
			Estimate Std. Error z value Pr(>|z|)    
			(Intercept)  0.37569    0.05416   6.937 4.01e-12 ***
			countries20 -0.64835    0.05388 -12.033  < 2e-16 ***
			countries80 -0.31199    0.05387  -5.792 6.97e-09 ***
			sanctions15 -0.13325    0.06208  -2.146  0.03183 *  
			sanctions20 -0.30356    0.06209  -4.889 1.01e-06 ***
			sanctions5   0.19186    0.06216   3.086  0.00203 ** 
			---
			Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
			(Dispersion parameter for binomial family taken to be 1)    
			
			Null deviance: 11783  on 8499  degrees of freedom
			Residual deviance: 11568  on 8494  degrees of freedom
			AIC: 11580
			Number of Fisher Scoring iterations: 4
		\end{verbatim}
	\begin{verbatim}
		> test2 # can reject knowledge 
		Analysis of Deviance Table
		Model 1: choice ~ 1
		Model 2: choice ~ countries + sanctions  
		Resid. Df Resid. Dev Df Deviance  Pr(>Chi)
		1      8499      11783                          
		2      8494      11568  5   215.15 < 2.2e-16 ***
		---
		Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
	\end{verbatim}
	\item The results of the additive model and the ANOVA suggest we can reject the null hypothesis and consider the Zscores of the prediction betas. All betas are statistically significant at the alpha = 0.05 level, holding all others constant. That suggests a change in the levels of the countries or sanctions variables will be associated with a high log odds of a change in the participant's choice, given we account for the other variable. Thus, a shift from 20 countries to 80 countries is expected to elicit a change in the log odds of choice as will a change in the number of sanctions from 5 to 20.
		
	\end{enumerate}
	
	\item
	If any of the explanatory variables are significant in this model, then:
	\begin{enumerate}
		\item
		For the policy in which nearly all countries participate [160 of 192], how does increasing sanctions from 5\% to 15\% change the odds that an individual will support the policy? (Interpretation of a coefficient)
			\begin{verbatim}
				For the policy in which nearly all countries participate:
				Yhat = 0.376 + -0.648(0 or 1) - 0.312(0 or 1) - 0.133(0 or 1)
				 - 0.304(0 or 1) + 0.192(0 or 1) 
				Yhat15percent = 0.376 + 0.192(1) = 0.568
				Yhat5percent = 0.376 - 0.133(1) = 0.243
				prob = odds / (1 + odds)
				= -0.325 / (1 - 0.325) = 0.4814815 decrease in probability
			\end{verbatim} 
%		\item
%		For the policy in which very few countries participate [20 of 192], how does increasing sanctions from 5\% to 15\% change the odds that an individual will support the policy? (Interpretation of a coefficient)
		\item
		What is the estimated probability that an individual will support a policy if there are 80 of 192 countries participating with no sanctions? 
		\begin{verbatim}
			Y_hat = 0.376 - 0.312(1) = 0.064
			0.064 / (1 + 0.064) = 0.06015038 probability 
		\end{verbatim}
		\item
		Would the answers to 2a and 2b potentially change if we included the interaction term in this model? Why? 
			\item[] Potentially, yes. Adding an interaction term adds another beta coefficient and potentially affects the shape of the model. 
		\begin{itemize}
			\item Perform a test to see if including an interaction is appropriate.
			\lstinputlisting[firstline=91, lastline=96]{PS2.R}
			\item[] The ANOVA suggests that there is no statistically significant difference between the two models, so we fail to reject H0 that including an interaction would have no effect on the model outcomes. 
		\end{itemize}
	\end{enumerate}
	\end{enumerate}


\end{document}
